---
title: "p8105_hw2_yl3181"
author: "Yuelin Long"
date: "September 28, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

Read and Clean the data. Generate a summary of the data.
```{r}
library(tidyverse)
nyc_subway = read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
             janitor::clean_names() %>%
             select(line:entry,vending, ada) %>% 
             mutate(entry = recode(entry, "YES" = as.logical(1), "NO" = as.logical(0)))
```

```{r}
nyc_subway_clean =  gather(nyc_subway, key = route_num, value = route_val, starts_with("route")) %>%
             separate(route_num, into = c("extra_route","route"), sep = 5) %>%
             select(-extra_route) %>%
             filter(!is.na(route_val)) %>%
             distinct(nyc_subway, line, station_name, .keep_all = TRUE)
```

Summary:

+ Data name: nyc_subway
+ Variables included: `r names(nyc_subway)`
+ Data cleaning steps: clean names -> select variables needed -> change character variables to logical -> gather route variables together -> clean route_num variable
+ Data dimension (rows x columns): `r dim(nyc_subway)`
+ Data quality: not tidy because the route1 - route11 should be values instead of column names 


Answer to the questions:

+ Number of distinct stations: `r nrow(nyc_subway_clean)`
+ Number of stations with ADA compliance: `r nrow(nyc_subway_clean[nyc_subway_clean$ada == 1,])`
+ Proportion of station entrance without vending: `r nrow(nyc_subway[(nyc_subway$vending == "NO")&(nyc_subway$entry == "TRUE"),]) /  nrow(nyc_subway[nyc_subway$vending == "NO",])`
+ Number of distinct stations serving A train: `r nrow(nyc_subway_clean[nyc_subway_clean$route_val == "A",])`
+ Number of distinct stations serving A train with ADA compliance: `r nrow(nyc_subway_clean[nyc_subway_clean$route_val == "A" & nyc_subway_clean$ada == "TRUE",])`


## Problem 2

Load the datasets:
```{r}
mr_trash = readxl::read_excel("HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N258") %>%
    janitor::clean_names() %>%
    filter(!is.na(dumpster)) %>%
    mutate(sports_balls = as.integer(sports_balls))

trash_2016 = readxl::read_excel("HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2016 Precipitation", range = "A2:B14") %>%
    janitor::clean_names() %>%
    mutate(year = 2016) %>%
    filter(!is.na(total))

trash_2017 = readxl::read_excel("HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>%
    janitor::clean_names() %>%
    mutate(year = 2017) %>%
    filter(!is.na(total))

library(plyr)

trash_2016_2017 = bind_rows(trash_2016, trash_2017)  %>%
    janitor::clean_names() %>%
    dplyr::rename(precipitation = total) %>%
    mutate(month = mapvalues(month,from = c(1:12), to = month.name))
```

Data summary:

+ The total N for Mr. Trash is `r nrow(mr_trash)`, while the examples of key variables are `r names(mr_trash[,1:7])`
+ The total N for trash in 2016 and 2017 is `r nrow(trash_2016_2017)` , while the key variables are `r names(trash_2016_2017)`
+ The total precipitation in 2017 is `r sum(trash_2017$total)`
+ The median number of sports balls in a dumpster in 2016 is `r median(mr_trash[mr_trash$year==2016,]$sports_balls, na.rm = TRUE)`


## Problem 3
```{r}
library(p8105.datasets)
brfss = janitor::clean_names(brfss_smart2010) %>%
    filter(topic == "Overall Health") %>%
    select(year:locationdesc, response, data_value) %>%
    spread(key = response, value = data_value) %>%
    mutate(excellent_verygood = Excellent + `Very good`) 
```

Data Summary:

+ Number of unique locations is `r nrow(distinct(brfss,locationdesc))`
+ Is every state represented: `r as.logical(nrow(distinct(brfss,locationabbr))>=50)`
+ State that is observed the most: `r dplyr::count(brfss,locationabbr,sort = TRUE)[1,1]`
+ In 2002, the median of the “Excellent” response value is `r median(brfss[brfss$year==2002,]$Excellent, na.rm = TRUE)`

Data Visualization:

A histogram of “Excellent” response values in the year 2002

```{r}
ggplot(brfss,aes(x = Excellent)) + geom_histogram(binwidth = 3)
```

A scatterplot showing the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010.

```{r}
ggplot(brfss[brfss$locationdesc %in% c("NY - New York County", "NY - Queens County"),],aes(x = year, y = Excellent, color = locationdesc)) + 
    geom_point() +
    geom_smooth(se = FALSE) 
```
